---
title: "Financial Advisor Talent Pools [DRAFT]"
author: 
- "Jacob Fischer, Branch Insights"
- "Matthew Milunski, Talent Management"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    highlight: kate
---


```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=FALSE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

```{r, echo=F, include=F, comment=NA}
# load packages ----
lapply(c("tidyverse", "data.table", "plotly"), function(package.name) {
    if (!is.element(package.name, installed.packages()[, 1])) {
    show(paste0("installing...", package.name))
    install.packages(package.name, dependencies = TRUE)
    library(package.name, character.only = TRUE)
    } else {
    library(package.name, character.only = TRUE)
    }
})

# load data ----
ksa <- fread(paste0(getwd(), "/ksa_df.csv"), header = TRUE, 
             stringsAsFactors = FALSE)
```

## Background
[PROVIDE SOME BACKGROUND]

## The Data
The Occupational Information Network known as [O*NET](https://www.onetcenter.org) stores various data on occupational definitions that helps people understand the United States' workforce. Select data categories from O*NET were used for analysis:

- Knowledge
- Skills
- Abilities
- Work Values
- Work Styles
- Interests

The resulting dataset contained `r nrow(ksa)` occupations and `r ncol(ksa)` variables. None of the variables were weighted for importance in this analysis.

## K-Means
```{r}
# k-means with 75 clusters ----
ksa_scaled_vars <- ksa %>%
    ## select for only the KSAs, etc
    select_if(is.numeric) %>%
    ## scale data
    mutate_all(scale)
set.seed(100)
kmeans_fit <- kmeans(ksa_scaled_vars, 75)
ksa_complete <- bind_cols(ksa,
                          data.frame(assigned_cluster = kmeans_fit$cluster))

# get cluster value that FAs belong to ----
fa_cluster_value <-
    ksa_complete$assigned_cluster[ksa_complete$title == "Personal Financial Advisors"]
similar_fa_jobs <- ksa_complete %>%
    filter(assigned_cluster == fa_cluster_value)

# make table ----
similar_df <- similar_fa_jobs %>%
    select(o_net_soc_code, Occupation = title) %>%
    filter(Occupation != "Personal Financial Advisors")
similar_df %>%
    knitr::kable("html", caption = "Jobs Related to Financial Advisors")
```
K-means clustering yielded `r nrow(similar_df)` occupations based on the previously mentioned categories.  

## Occupations Closest to Financial Advisor
From the previous cluster solution, the distances were calculated within that cluster in order to find which jobs were most similar to the Financial Advisor. Those with a "Similarity to FA" value closest to zero are more like the financial advisor in their roles.
```{r}
# create dataframe from `similar_fa_jobs` cluster and rescale ----
fa_jobs_scaled <- similar_fa_jobs %>%
    select(administration_and_management:social) %>%
    mutate_all(scale)

# create matrix of distances and attach to `similar_fa_jobs`----
e_dist <- as.matrix(dist(fa_jobs_scaled))
similar_fa_jobs2 <- data.frame(similar_fa_jobs,
                               euclidean_dist = e_dist[, 10]) %>%
    select(o_net_soc_code, title, euclidean_dist) %>%
    arrange(euclidean_dist)

# make a table ----
similar_fa_jobs2 %>% select(Occupation = title,
                            `Similarity to FA` = euclidean_dist) %>%
    knitr::kable("html", title = "Occupations Ranked as Most Similar to Financial Advisors")
```

```{r all jobs, include=F}
# all jobs ----
onet.dist <- function(data, job_title) {
    ## get KSAs, etc from data
    all_jobs <- data[, 4:ncol(data)]
    ## scale, create distance matrix, and format as matrix
    scaled_dist <- as.matrix(dist(scale(all_jobs)))
    ## join job title and other info to matrix
    dist_df <- data.frame(data[, 1:3], scaled_dist)
    ## find column index that has zero distance relative to the job_title
    equals_zero <-
        as.data.frame(which(dist_df == 0, arr.ind = TRUE))
    row_ref <- which(dist_df$title == job_title)
    col_ref <- equals_zero$col[row_ref]
    ## return basic job info and associated distances
    all_jobs2 <- subset(dist_df,
                        select = c(o_net_soc_code, title, description,
                                   col_ref))
    return(all_jobs2)
}
onet.dist(ksa, "Personal Financial Advisors")
```

## Jobs when weighting certain relationship-based variables

## Recommendations
[LIST RECOMMENDATIONS]


