---
title: "Financial Advisor Talent Pools"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  rmdformats::html_clean:
    highlight: kate
---


```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=FALSE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

```{r, echo=F, include=F, comment=NA}
# load packages ----
lapply(c("tidyverse", "plotly"), function(package.name) {
    if (!is.element(package.name, installed.packages()[, 1])) {
    show(paste0("installing...", package.name))
    install.packages(package.name, dependencies = TRUE)
    library(package.name, character.only = TRUE)
    } else {
    library(package.name, character.only = TRUE)
    }
})

# load data ----
ksa <- read.csv(paste0(getwd(), "/ksa_df.csv"), header = TRUE, 
                stringsAsFactors = FALSE)
```

## Background
This analysis is intended to identify similar jobs for the financial advisor using data from the Occupational Information Network ([O*NET](https://www.onetcenter.org)). Various solutions are presented.

## The Data
O\*NET stores various data on occupational definitions that helps people understand the United States' workforce. Select data categories from O*NET were used for analysis:

- Knowledge
- Skills
- Abilities
- Work Values
- Work Styles
- Interests

### Data Preprocessing
*Knowledge, skills, abilities, work styles.* The data from the knowledge, skills, abilities, and work styles were preprocessed in a similar fashion. The importance measures for each dataset were used (`scale_id == 'IM'`). Observations that were recommended for suppression (`recommend_suppress`) or not relevant (`not_relevant`) were removed from the dataset.  

*Work values*. The work values data used the *mean extent rating* for each work value.  

*Interests*. The interests data used the RIASEC level of each interest (`scale_id == 'OI'`).  

The resulting dataset contained `r nrow(ksa)` occupations and `r ncol(ksa)` variables. Please refer to the [O*NET 22.0 Database Data Dictionary](https://www.onetcenter.org/dictionary/22.0/excel/) for more information on the data used in this analysis.  

## K-Means
```{r}
# k-means with 75 clusters ----
ksa_scaled_vars <- ksa %>%
    ## select for only the KSAs, etc
    select_if(is.numeric) %>%
    ## scale data
    mutate_all(scale)
set.seed(100)
kmeans_fit <- kmeans(ksa_scaled_vars, 75)
ksa_complete <- bind_cols(ksa,
                          data.frame(assigned_cluster = kmeans_fit$cluster))

# get cluster value that FAs belong to ----
fa_cluster_value <-
    ksa_complete$assigned_cluster[ksa_complete$title == "Personal Financial Advisors"]
similar_fa_jobs <- ksa_complete %>%
    filter(assigned_cluster == fa_cluster_value)

# make table ----
similar_df <- similar_fa_jobs %>%
    select(o_net_soc_code, Occupation = title) %>%
    filter(Occupation != "Personal Financial Advisors")
```
K-means clustering was used to segment the occupations into similar groups based upon all numeric variables in the data. No variables were weighted. The algorithm was instructed to find a *k*=75 solution since it provides a manageable number of occupations. The solution presented `r nrow(similar_df)` jobs with similar characteristics to the personal financial advisor.  

```{r}
similar_df %>%
    knitr::kable("html", caption = "Jobs Related to Financial Advisors")
```

## Occupations Closest to Financial Advisor
From the previous cluster solution, the distances were calculated within that cluster in order to find which jobs were most similar to the Financial Advisor. Those with a "Similarity to FA" value closest to zero are more like the financial advisor in their roles.
```{r}
# create dataframe from `similar_fa_jobs` cluster and rescale ----
fa_jobs_scaled <- similar_fa_jobs %>%
    select(administration_and_management:social) %>%
    mutate_all(scale)

# create matrix of distances and attach to `similar_fa_jobs`----
e_dist <- as.matrix(dist(fa_jobs_scaled))
similar_fa_jobs2 <- data.frame(similar_fa_jobs,
                               euclidean_dist = e_dist[, 10]) %>%
    select(o_net_soc_code, title, euclidean_dist) %>%
    arrange(euclidean_dist)

# make a table ----
similar_fa_jobs2 %>% select(Occupation = title,
                            `Similarity to FA` = euclidean_dist) %>%
    knitr::kable("html", title = "Occupations Ranked as Most Similar to Financial Advisors")
```


```{r all jobs, include=F}
# all jobs ----
onet.dist <- function(data, job_title, show_topn = 5) {
    ## get KSAs, etc from data
    all_jobs <- data[, 4:ncol(data)]
    ## scale, create distance matrix, and format as matrix
    scaled_dist <- as.matrix(dist(scale(all_jobs)))
    ## join job title and other info to matrix
    dist_df <- data.frame(data[, 1:2], scaled_dist)
    ## find column index that has zero distance relative to the job_title
    equals_zero <-
        as.data.frame(which(dist_df == 0, arr.ind = TRUE))
    row_ref <- which(dist_df$title == job_title)
    col_ref <- equals_zero$col[row_ref]
    ## retain basic job info and associated distances
    all_jobs2 <- subset(dist_df,
                        select = c(o_net_soc_code, title, col_ref))
    all_jobs2 <- all_jobs2[order(all_jobs2[, 3]), ]
    ## return top n-rows
    top_jobs <- all_jobs2[1:show_topn, ]
    return(top_jobs)
}
onet.dist(ksa, "Personal Financial Advisors", 20)
```

## Weighting More Interpersonal Skill Variables
The related jobs shown in the previous sections were the result of distances calculated from unweighted variables. That is, all KSAOs were considered equal to each other. Weighting certain interpersonal interaction variables may help create a revised list where certain social abilities are given more emphasis in the distance calculation.  

Nine different qualities were selected for weighting based upon researcher knowledge of the company-specific FA role:  

- Relationships
- Concern for others
- Social orientation
- Social
- Persuasion
- Active listening
- Service orientation
- Sales and marketing
- Customer and personal service  

The above qualities were weighted by a factor of two (2). This seemed to create a list that removed more analyst-based roles and used roles with greater "people skill" involvement.
```{r soft skill weights}
# scale data ----
scaled <- ksa %>% 
    mutate_at(vars(administration_and_management:ncol(.)), scale) 

# weight the variables by a factor of 10 ----
scaled_vars <- c("relationships", "concern_for_others", 
                "social_orientation", "social", "persuasion",
                "active_listening", "service_orientation",
                "sales_and_marketing", "customer_and_personal_service")
scaled[, scaled_vars] <- lapply(scaled[, scaled_vars], function(x) {
    ifelse(scaled$o_net_soc_code == "13-2052.00", x * 2, x)
})

# redo the distances ----
weighted_distances <- scaled %>% 
    select(administration_and_management:ncol(.)) %>% 
    dist() %>% 
    as.matrix %>%
    data.frame(select(scaled, o_net_soc_code:description), .)

# get the table ----
weighted_tbl <- weighted_distances %>% 
    select(o_net_soc_code, title, distance = X96) %>% 
    arrange(distance) %>% 
    top_n(-21, distance) 

weighted_tbl %>% 
    select(title, distance) %>% 
    knitr::kable("html", title = "Occupations Similar to FA Weighted for Interpersonal Skills")
```

## Conclusion
This analysis provided potential occupations similar to financial advisors through finding commonalities in knowledge, skills, abilities, work values, work styles, and interests. K-means clustering, unweighted euclidean distances, and weighted euclidean distances were used to find similar occupations. The weighted distance analysis was selected for stakeholder presentation as the occupations overlapped with prior information while providing novel occupations to consider. Future research should test the validity of these selected jobs in predicting FA performance.  




